---
title: "Assignment1"
author: "Study Group 10"
date: "12/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(readxl)
library(dplyr)
library(tidyverse)
library(Hmisc)
```


## Empirical Finance Assignment 1

This assignment gets you started working with prices and returns. You should do everything in R even though we are only working with six different stocks and it is possible to do some things manually. If you have a sample of hundreds of stocks this might become more difficult.

1. Go to CANVAS and download the data for PS1 Monthly.xlsx. The data was downloaded from the
CRSP (The Center for Research in Security Prices) data base via WRDS (Wharton Research Service)
that is available for LBS students. Note the file contains data about Microsoft, ExxonMobil (previously
Exxon), General Electric, JP Morgan Chase (previously Chemical Banking and Chase Manhattan),
Intel, Citigroup (previously Primerica and Travellers Group). In addition, the columns vwretd (ewretd)
and vwredx (ewretx) contain value-weighted (equal-weighted) total returns and total returns excluding
dividends for the CRSP index that contains stocks from NYSE, AMEX, and NASDAQ. Finally, sprtrn
contains the total return for the S&P 500 Composite Index.

## Data Investigation

```{r load data}
# loading the data
data <- read_excel("PS1_Daily.xlsx")
janitor::clean_names(data)

# investigating the data
glimpse(data) 
summary(data)
head(data, 5)
describe(data)
```


## Part II
2. Make sure you understand how the holding period returns (RET) are calculated
```{r}
```


3. Use the holding period returns to create a total return index for the MSFT and GE stocks and the S&P
500 index, which shows the theoretical growth in value of an investment in the stock assuming that
dividends are reinvested (normalize the start value to 1). Do the same for the returns that abstract
from dividend payments (i.e., use RETX instead). Plot the investments with and without dividends
for each stock separately. How do dividends affect the results stock by stock?
```{r}

```



4. Normalize the price of GE using the adjusted number of shares. Plot the adjusted price against the
unadjusted prices. Discuss.
```{r}

```



5. The holding period returns are normal returns. Generate a new variable that contains the corresponding
log returns (LRET). Calculate the mean, variance, skewness, and kurtosis of the normal and the log
returns. Plot the normal against the log returns for MSFT. Briefly discuss your results!
```{r}

```



6. Go to CANVAS and download the data for PS1 Daily.xlsx. This file contains two worksheets.
HPR Daily contains the daily holding period returns for the six stocks, the S&P 500 Composite Index
and the value-weighted market portfolio (including dividends) from CRSP. Prices Daily contains the
prices for the six stocks and the S&P 500 Composite Index.
```{r}

```



7. Construct a daily total return index for MSFT and GE stocks and the S&P 500 index and plot them
against each other. Compare your results with the monthly total return indices from above. Are there
any differences? Discuss.
```{r}

```



8. As before, the holding period returns are normal returns. Create log returns. Calculate the mean,
variance, skewness, and kurtosis of the normal and log returns. Compare and discuss your results with
the results from monthly frequency.
```{r}

```


9. Compare the statistical properties of the log holding period return time series both for monthly and
daily returns. Plot a histogram and discuss how the empirical distributions relate to the normal
distribution.
```{r}

```


10. Pick three stocks and the S&P 500 index (either you can use MSFT, GE and JPM or adapt the code to
pick three random stocks). You will need the holding period returns (both normal and log returns)
and the total return indices you created.
```{r}

```


11. Calculate the covariance matrix for the log return series, using both the returns and returns squared.
Discuss your results briefly.
```{r}

```



12. Plot the ACF (autocorrelation function) for prices, returns, returns squared, and absolute returns.
Discuss the results!
```{r}

```



13. Use the three assets and make up a portfolio by assigning arbitrary portfolio weights. What does it
imply if you keep the weights fixed over time?
```{r}

```



14. Calculate the portfolio returns and use them to calculate the evolution of a $ 1 investment in the
portfolio over the sample period. Plot the result against the evolution of a $ 1 investment in each of
the three stocks. Discuss the result.
```{r}

```



15. Portfolio theory with matrix algebra: 
(see assignment)
```{r}

```
















